---
title: "IQVIA21_analysis_w/R"
format: html
editor: visual
---

```{r}
library(dplyr)
library(readr)
library(ggplot2)
library(tidyverse)
```

# !!import data

```{r}
patient_demographics <- read_csv("Q.patient_demographics.csv")

usage <- read_csv("Q.usage.csv")
usage <- usage %>%
  filter_at(vars(antibiotic_name, dosage_form, provider_description), all_vars(!is.na(.)))

unfiltered <- read_csv("Q.unfiltered100k_1.csv")
unfiltered <- unfiltered
  #na.omit()

#*add age group___
patient_demographics <- patient_demographics %>%
  filter (pat_age_nbr >=0) %>%
  mutate(age_cat = case_when(
  pat_age_nbr < 21 ~ "Pediatrics",
  pat_age_nbr >= 65 ~ "Elders",
  TRUE ~ "Adults"
  ))

unfiltered <- unfiltered %>%
filter (pat_age_nbr >=0) %>%
  mutate(age_cat = case_when(
  pat_age_nbr < 21 ~ "Pediatrics",
  pat_age_nbr >= 65 ~ "Elders",
  TRUE ~ "Adults"
  ))

###___

RX <- unfiltered %>%
  select(usc_lvl3_desc,
    unifm_prod_nm,
    pat_age_nbr,
    rx_dosage_amt,
    dosage_form_nm,
    strnt_desc,
    pri_spcl_desc,
    provider_st_cd,
    age_cat) %>%
  filter(rx_dosage_amt > 0)

copayment <- unfiltered %>%
  select(usc_lvl3_desc,
         unifm_prod_nm,
      pri_spcl_desc,
      pat_age_nbr,
      copay_amt,
      cust_prc,
      pat_pay_amt,
      model_typ_cd,
      model_typ_nm,
      age_cat)%>%
  filter(copay_amt > 0)
```

#Patient demographics

##draw graph by antibiotic category

```{r}
ggplot(patient_demographics, aes(x=age_cat))+
  geom_bar()+
  facet_wrap(vars(usc_lvl3_desc))
```

## !!make top used antibiotic groups within each category

```{r}
top_prod <- patient_demographics %>%
  filter(usc_lvl3_desc == "ANTNFCT SYSTEMIC, OTHER") %>%
  group_by(unifm_prod_nm) %>%
  summarise(count = sum(n)) %>%
  arrange(desc(count)) %>%
  slice_head(n=5) 
antnfct_systemic<-top_prod$unifm_prod_nm



top_prod <- patient_demographics %>%
  filter(usc_lvl3_desc == "ANTNFCT, BROAD/MEDIUM SPECTRUM") %>%
  group_by(unifm_prod_nm) %>%
  summarise(count = sum(n)) %>%
  arrange(desc(count)) %>%
  slice_head(n=5) 
antnfct_broad<-top_prod$unifm_prod_nm


top_prod <- patient_demographics %>%
  filter(usc_lvl3_desc == "B-LACTAM, INCREASED ACTIVITY") %>%
  group_by(unifm_prod_nm) %>%
  summarise(count = sum(n)) %>%
  arrange(desc(count)) %>%
  slice_head(n=5) 
b_lactam<-top_prod$unifm_prod_nm


top_prod <- patient_demographics %>%
  filter(usc_lvl3_desc == "OPHTHALMIC ANTI-INFECTIVES") %>%
  group_by(unifm_prod_nm) %>%
  summarise(count = sum(n)) %>%
  arrange(desc(count)) %>%
  slice_head(n=5) 
ophthalmic<-top_prod$unifm_prod_nm


top_prod <- patient_demographics %>%
  filter(usc_lvl3_desc == "OTIC ANTI-INFECTIVES") %>%
  group_by(unifm_prod_nm) %>%
  summarise(count = sum(n)) %>%
  arrange(desc(count)) %>%
  slice_head(n=5) 
otic<-top_prod$unifm_prod_nm


top_prod <- patient_demographics %>%
  filter(usc_lvl3_desc == "OXAZOLIDINONES") %>%
  group_by(unifm_prod_nm) %>%
  summarise(count = sum(n)) %>%
  arrange(desc(count)) %>%
  slice_head(n=5) 
oxazolidinones<-top_prod$unifm_prod_nm


top_prod <- patient_demographics %>%
  filter(usc_lvl3_desc == "SULFONAMIDES & TRIMETHOPRIM") %>%
  group_by(unifm_prod_nm) %>%
  summarise(count = sum(n)) %>%
  arrange(desc(count)) %>%
  slice_head(n=5) 
sulf_trim<-top_prod$unifm_prod_nm


top_prod <- patient_demographics %>%
  filter(usc_lvl3_desc == "VAGINAL ANTI-INFECTIVES") %>%
  group_by(unifm_prod_nm) %>%
  summarise(count = sum(n)) %>%
  arrange(desc(count)) %>%
  slice_head(n=5) 
vaginal<-top_prod$unifm_prod_nm
```

##draw graph by antibiotics

```{r}
patient_demographics %>%
  filter(unifm_prod_nm == antnfct_systemic) %>%
  group_by(unifm_prod_nm) %>%
ggplot(aes(x=age_cat, y=sum(n)))+
  geom_col()+
  facet_wrap(vars(unifm_prod_nm))

patient_demographics %>%
  filter(unifm_prod_nm == antnfct_broad) %>%
  group_by(unifm_prod_nm) %>%
ggplot(aes(x=age_cat, y=sum(n)))+
  geom_col()+
  facet_wrap(vars(unifm_prod_nm))

patient_demographics %>%
  filter(unifm_prod_nm == b_lactam) %>%
  group_by(unifm_prod_nm) %>%
ggplot(aes(x=age_cat, y=sum(n)))+
  geom_col()+
  facet_wrap(vars(unifm_prod_nm))

patient_demographics %>%
  filter(unifm_prod_nm == ophthalmic) %>%
  group_by(unifm_prod_nm) %>%
ggplot(aes(x=age_cat, y=sum(n)))+
  geom_col()+
  facet_wrap(vars(unifm_prod_nm))

patient_demographics %>%
  filter(unifm_prod_nm == otic) %>%
  group_by(unifm_prod_nm) %>%
ggplot(aes(x=age_cat, y=sum(n)))+
  geom_col()+
  facet_wrap(vars(unifm_prod_nm))

patient_demographics %>%
  filter(unifm_prod_nm == oxazolidinones) %>%
  group_by(unifm_prod_nm) %>%
ggplot(aes(x=age_cat, y=sum(n)))+
  geom_col()+
  facet_wrap(vars(unifm_prod_nm))

patient_demographics %>%
  filter(unifm_prod_nm == sulf_trim) %>%
  group_by(unifm_prod_nm) %>%
ggplot(aes(x=age_cat, y=sum(n)))+
  geom_col()+
  facet_wrap(vars(unifm_prod_nm))

patient_demographics %>%
  filter(unifm_prod_nm == vaginal) %>%
  group_by(unifm_prod_nm) %>%
ggplot(aes(x=age_cat, y=sum(n)))+
  geom_col()+
  facet_wrap(vars(unifm_prod_nm))
```

#Usage

##trial plots for novel visualization formats

```{r}
#hide
#Faceted Heatmap by Dosage Form
#helped by AI
usage %>%
  filter(antibiotic_name == antnfct_systemic)%>%
ggplot(aes(provider_description,
           antibiotic_name,
           fill = count)) +
  geom_tile() +
  facet_wrap(~ dosage_form) +
  scale_fill_gradient(low = "yellow", high = "red")+
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 90)
  )
```

```{r}
#hide
#bubble
usage %>%
  filter(antibiotic_name == antnfct_systemic)%>%
ggplot(aes(provider_description,
           antibiotic_name,
           size = count,
           color = dosage_form)) +
  geom_point(alpha = .7) +
  scale_size(range = c(3, 9)) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 90)
  )
```

##draw bar graphs

```{r}
#bar
usage %>%
  filter(antibiotic_name  %in%  antnfct_systemic,
         count >10)%>%
ggplot(aes(x = provider_description,
           y = count,
           fill = dosage_form)) +
  geom_col(position = "stack") +
  coord_flip()+
  geom_text(aes(label = count), vjust = -0.2)+
  #facet_wrap(~ antibiotic_name)+
  labs(title = "Distribution of Prescriptions by Provider and Dosage",
    subtitle = "ANTNFCT SYSTEMIC, OTHER")

```

```{r}
usage %>%
  filter(antibiotic_name  %in%  antnfct_broad,
         count >10)%>%
  group_by(antibiotic_name) %>%
  slice_max(n=10,
            order_by = count) %>%
ggplot(aes(x = provider_description,
           y = count,
           fill = dosage_form)) +
  geom_col(position = "stack") +
  coord_flip() +
  facet_wrap(~ antibiotic_name) + 
  geom_text_repel(
    aes(label = count),
    position = position_stack(vjust = 0.3),
    size = 3)+
  labs(title = "Distribution of Prescriptions by Provider and Dosage",
    subtitle ="ANTNFCT, BROAD/MEDIUM SPECTRUM")

#if i stack all antibiotics together and only focus on the condition:
usage %>%
  filter(antibiotic_name  %in%  antnfct_broad,
         count >10)%>%
  group_by(antibiotic_name) %>%
  slice_max(n=10,
            order_by = count) %>%
ggplot(aes(x = provider_description,
           y = count,
           fill = dosage_form)) +
  geom_col(position = "stack") +
  coord_flip() +
  labs(title = "Distribution of Prescriptions by Provider and Dosage",
    subtitle ="ANTNFCT, BROAD/MEDIUM SPECTRUM")


```

```{r}
usage %>%
  filter(antibiotic_name  %in%  b_lactam,
         count >10)%>%
  group_by(antibiotic_name) %>%
  slice_max(n=10,
            order_by = count) %>%
  ggplot(aes(x = provider_description,
           y = count,
           fill = dosage_form)) +
  geom_col(position = "stack") +
  coord_flip()+
  facet_wrap(~ antibiotic_name)+
  geom_text(
    aes(label = count),
    position = position_stack(vjust = 0.5),
    size = 3)+
  labs(title = "Distribution of Prescriptions by Provider and Dosage",
    subtitle ="B-LACTAM, INCREASED ACTIVITY")


usage %>%
  filter(antibiotic_name  %in%  ophthalmic,
         count >10)%>%
  group_by(antibiotic_name) %>%
  slice_max(n=10,
            order_by = count) %>%
  ggplot(aes(x = provider_description,
           y = count,
           fill = dosage_form)) +
  geom_col(position = "stack") +
  coord_flip()+
  facet_wrap(~ antibiotic_name) +
  geom_text(
    aes(label = count),
    position = position_stack(vjust = 0.5),
    size = 3)+
  labs(title = "Distribution of Prescriptions by Provider and Dosage",
    subtitle ="OPHTHALMIC ANTI-INFECTIVES")


usage %>%
  filter(antibiotic_name  %in%  otic,
         count >10)%>%
  group_by(antibiotic_name) %>%
  slice_max(n=10,
            order_by = count) %>%
  ggplot(aes(x = provider_description,
           y = count,
           fill = dosage_form)) +
  geom_col(position = "stack") +
  coord_flip()+
  facet_wrap(~ antibiotic_name)+
  geom_text(
    aes(label = count),
    position = position_stack(vjust = 0.5),
    size = 3)+
  labs(title = "Distribution of Prescriptions by Provider and Dosage",
    subtitle = "OTIC ANTI-INFECTIVES")


usage %>%
  filter(antibiotic_name  %in%  oxazolidinones,
         count >10)%>%
  group_by(antibiotic_name) %>%
  slice_max(n=10,
            order_by = count) %>%
  ggplot(aes(x = provider_description,
           y = count,
           fill = dosage_form)) +
  geom_col(position = "stack") +
  coord_flip()+
  facet_wrap(~ antibiotic_name)+
  geom_text(
    aes(label = count),
    position = position_stack(vjust = 0.5),
    size = 3) +
  labs(title = "Distribution of Prescriptions by Provider and Dosage",
    subtitle = "OXAZOLIDINONES")


usage %>%
  filter(antibiotic_name  %in%  sulf_trim,
         count >10)%>%
  group_by(antibiotic_name) %>%
  slice_max(n=10,
            order_by = count) %>%
  ggplot(aes(x = provider_description,
           y = count,
           fill = dosage_form)) +
  geom_col(position = "stack") +
  coord_flip()+
  facet_wrap(~ antibiotic_name)+
  geom_text(
    aes(label = count),
    position = position_stack(vjust = 0.5),
    size = 3) +
  labs(title = "Distribution of Prescriptions by Provider and Dosage",
    subtitle = "SULFONAMIDES & TRIMETHOPRIM")


usage %>%
  filter(antibiotic_name  %in%  vaginal,
         count >10)%>%
  group_by(antibiotic_name) %>%
  slice_max(n=10,
            order_by = count) %>%
  ggplot(aes(x = provider_description,
           y = count,
           fill = dosage_form)) +
  geom_col(position = "stack") +
  coord_flip()+
  facet_wrap(~ antibiotic_name) +
  geom_text_repel(
    aes(label = count),
    position = position_stack(vjust = 0.5),
    size = 3) +
  labs(title = "Distribution of Prescriptions by Provider and Dosage",
    subtitle = "VAGINAL ANTI-INFECTIVES")
```

#RX

##boxplots of the distribution of RX amount (range) 1) by antibiotics; 2) across prescribing conditions, by antibiotic categories; 3) acrosss age groups

```{r}
RX %>%
filter(usc_lvl3_desc == "B-LACTAM, INCREASED ACTIVITY") %>%
  group_by(unifm_prod_nm) %>%
  ggplot(aes(x=unifm_prod_nm, y=rx_dosage_amt))+
  geom_violin(scale = "width")+
  geom_boxplot(width=0.2,
               color = "blue")+
  theme(axis.text=element_text(size=7),
        strip.text = element_text(size = 7)) +
  labs(title = "Distribution of RX amount by antibiotic",
       subtitle = "B-LACTAM, INCREASED ACTIVITY",
       x = "Antibiotics",
       y = "Total Rx Amount")

RX %>%
filter(usc_lvl3_desc == "B-LACTAM, INCREASED ACTIVITY") %>%
  group_by(pri_spcl_desc) %>%
  mutate(num = n()) %>%
  ungroup() %>%
  filter(num >5) %>%
  ggplot(aes(x=unifm_prod_nm, y=rx_dosage_amt, color=dosage_form_nm))+
  geom_boxplot(width=0.5)+
  facet_wrap(~ pri_spcl_desc)+
  theme(axis.text=element_text(size=7),
        strip.text = element_text(size = 7)) +
  labs(title = "Distribution of RX amount across prescribing conditions",
       subtitle = "B-LACTAM, INCREASED ACTIVITY")

RX %>%
filter(usc_lvl3_desc == "B-LACTAM, INCREASED ACTIVITY") %>%
  group_by(age_cat) %>%
  mutate(indivdual = n()) %>%
  ungroup() %>%
  filter(indivdual >5) %>%
  ggplot(aes(x=age_cat, y=rx_dosage_amt, color=dosage_form_nm))+
  geom_boxplot(width=0.5)+
  facet_wrap(~ unifm_prod_nm)+
  theme(axis.text=element_text(size=7),
        strip.text = element_text(size = 7))+
  labs(title = "Distribution of RX amount across age groups",
       subtitle = "B-LACTAM, INCREASED ACTIVITY")
```

```{r}
RX %>%
filter(usc_lvl3_desc == "SULFONAMIDES & TRIMETHOPRIM",) %>%
  group_by(unifm_prod_nm) %>%
  filter(n() >= 5)%>%
  ggplot(aes(x=unifm_prod_nm, y=rx_dosage_amt))+
  geom_violin(scale = "width")+
  geom_boxplot(width=0.2,
               aes(color = unifm_prod_nm))+
  theme(axis.text=element_text(size=7),
        strip.text = element_text(size = 7)) +
  labs(title = "Distribution of RX amount by antibiotic",
       subtitle = "SULFONAMIDES & TRIMETHOPRIM",
       color = "Antibiotics",
       x = "Antibiotics",
       y = "Total Rx Amount")

RX %>%
filter(usc_lvl3_desc == "SULFONAMIDES & TRIMETHOPRIM") %>%
  group_by(pri_spcl_desc) %>%
  mutate(num = n()) %>%
  ungroup() %>%
  filter(num >5) %>%
  ggplot(aes(x=unifm_prod_nm, y=rx_dosage_amt, color=dosage_form_nm))+
  geom_boxplot(width=0.5)+
  facet_wrap(~ pri_spcl_desc)+
  theme(axis.text=element_text(size=7),
        strip.text = element_text(size = 7),
         axis.text.x = element_text(angle = 90))+
  labs(title = "Distribution of RX amount across prescribing conditions",
       subtitle = "SULFONAMIDES & TRIMETHOPRIM")


RX %>%
filter(usc_lvl3_desc == "SULFONAMIDES & TRIMETHOPRIM") %>%
  group_by(age_cat) %>%
  mutate(indivdual = n()) %>%
  ungroup() %>%
  filter(indivdual >3) %>%
  ggplot(aes(x=age_cat, y=rx_dosage_amt, color=dosage_form_nm))+
  geom_boxplot(width=0.5)+
  facet_wrap(~ unifm_prod_nm)+
  theme(axis.text=element_text(size=7),
        strip.text = element_text(size = 7))+
  labs(title = "Distribution of RX amount across age groups",
       subtitle = "SULFONAMIDES & TRIMETHOPRIM")
```

```{r}
RX %>%
filter(usc_lvl3_desc == "ANTNFCT, BROAD/MEDIUM SPECTRUM") %>%
  group_by(unifm_prod_nm) %>%
  filter(n() >= 50)%>%
  ggplot(aes(x=unifm_prod_nm, y=rx_dosage_amt))+
  geom_violin(scale = "width")+
  geom_boxplot(width=0.2,
               aes(color = unifm_prod_nm))+
  scale_y_log10()+
  theme(axis.text=element_text(size=7),
        strip.text = element_text(size = 7),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  labs(title = "Distribution of RX amount by antibiotic",
       subtitle = "ANTNFCT, BROAD/MEDIUM SPECTRUM",
       color = "Antibiotics",
       x = "Antibiotics",
       y = "Total Rx Amount")

#########

#the conditions for ANTNFCT, BROAD/MEDIUM SPECTRUM are a bit complex
RX %>%
filter(usc_lvl3_desc == "ANTNFCT, BROAD/MEDIUM SPECTRUM") %>%
  group_by(pri_spcl_desc) %>%
  mutate(num = n()) %>%
  ungroup() %>%
  filter(num >100) %>%
  ggplot(aes(x=unifm_prod_nm, y=rx_dosage_amt, color=dosage_form_nm))+
  geom_boxplot(width=0.5)+
  facet_wrap(~ pri_spcl_desc)+
  theme(axis.text=element_text(size=7),
        strip.text = element_text(size = 7))+
  labs(title = "Distribution of RX amount across prescribing conditions",
       subtitle = "ANTNFCT, BROAD/MEDIUM SPECTRUM")

#since the results are overwhelming, look into some conditions selectively
##reference list
RX %>%
filter(usc_lvl3_desc == "ANTNFCT, BROAD/MEDIUM SPECTRUM") %>%
  group_by(pri_spcl_desc) %>%
  mutate(num = n()) %>%
  ungroup() %>%
  filter(num >100)%>%
  distinct(pri_spcl_desc) %>%   
  print(n = Inf)    
##boxplot for a single condition
RX %>%
filter(usc_lvl3_desc == "ANTNFCT, BROAD/MEDIUM SPECTRUM",
       pri_spcl_desc == "EMERGENCY MEDICINE") %>%
  ggplot(aes(x=unifm_prod_nm, y=rx_dosage_amt, color=dosage_form_nm))+
  geom_boxplot(width=0.5)+
  facet_wrap(~ pri_spcl_desc)+
  theme(axis.text=element_text(size=7),
        strip.text = element_text(size = 7),
        axis.text.x = element_text(angle = 90))+
  labs(title = "Distribution of RX amount across prescribing conditions",
       subtitle = "ANTNFCT, BROAD/MEDIUM SPECTRUM")


#########

RX %>%
filter(usc_lvl3_desc == "ANTNFCT, BROAD/MEDIUM SPECTRUM") %>%
  group_by(age_cat) %>%
  mutate(indivdual = n()) %>%
  ungroup() %>%
  filter(indivdual >3) %>%
  ggplot(aes(x=age_cat, y=rx_dosage_amt, color=dosage_form_nm))+
  geom_boxplot(width=0.5)+
  facet_wrap(~ unifm_prod_nm)+
  theme(axis.text=element_text(size=7),
        strip.text = element_text(size = 7))+
  labs(title = "Distribution of RX amount across age groups",
       subtitle = "ANTNFCT, BROAD/MEDIUM SPECTRUM")

```

```{r}
RX %>%
filter(usc_lvl3_desc == "ANTNFCT SYSTEMIC, OTHER",) %>%
  group_by(unifm_prod_nm) %>%
  filter(n() >= 5)%>%
  ggplot(aes(x=unifm_prod_nm, y=rx_dosage_amt))+
  geom_violin(scale = "width")+
  geom_boxplot(width=0.2,
               aes(color = unifm_prod_nm))+
  scale_y_log10()+
  theme(axis.text=element_text(size=7),
        strip.text = element_text(size = 7),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  labs(title = "Distribution of RX amount by antibiotic",
       subtitle = "ANTNFCT SYSTEMIC, OTHER",
       color = "Antibiotics",
       x = "Antibiotics",
       y = "Total Rx Amount")



RX %>%
filter(usc_lvl3_desc == "ANTNFCT SYSTEMIC, OTHER") %>%
  group_by(pri_spcl_desc) %>%
  mutate(num = n()) %>%
  ungroup() %>%
  #filter(num >5) %>%
  ggplot(aes(x=unifm_prod_nm, y=rx_dosage_amt, color=dosage_form_nm))+
  geom_boxplot(width=0.5)+
  facet_wrap(~ pri_spcl_desc)+
  theme(axis.text=element_text(size=7),
        strip.text = element_text(size = 7))+
  labs(title = "Distribution of RX amount across prescribing conditions",
       subtitle = "ANTNFCT SYSTEMIC, OTHER")


RX %>%
filter(usc_lvl3_desc == "ANTNFCT SYSTEMIC, OTHER") %>%
  group_by(age_cat) %>%
  mutate(indivdual = n()) %>%
  ungroup() %>%
  filter(indivdual >3) %>%
  ggplot(aes(x=age_cat, y=rx_dosage_amt, color=dosage_form_nm))+
  geom_boxplot(width=0.5)+
  facet_wrap(~ unifm_prod_nm)+
  theme(axis.text=element_text(size=7),
        strip.text = element_text(size = 7))+
  labs(title = "Distribution of RX amount across age groups",
       subtitle = "ANTNFCT SYSTEMIC, OTHER")
```

```{r}
RX %>%
filter(usc_lvl3_desc == "OXAZOLIDINONES",) %>%
  group_by(unifm_prod_nm) %>%
  filter(n() >= 5)%>%
  ggplot(aes(x=unifm_prod_nm, y=rx_dosage_amt))+
  geom_violin(scale = "width")+
  geom_boxplot(width=0.2,
               aes(color = unifm_prod_nm))+
  scale_y_log10()+
  theme(axis.text=element_text(size=7),
        strip.text = element_text(size = 7),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  labs(title = "Distribution of RX amount by antibiotic",
       subtitle = "OXAZOLIDINONES",
       color = "Antibiotics",
       x = "Antibiotics",
       y = "Total Rx Amount")


RX %>%
filter(usc_lvl3_desc == "OXAZOLIDINONES") %>%
  group_by(pri_spcl_desc) %>%
  mutate(num = n()) %>%
  ungroup() %>%
  #filter(num >5) %>%
  ggplot(aes(x=unifm_prod_nm, y=rx_dosage_amt, color=dosage_form_nm))+
  geom_boxplot(width=0.5)+
  facet_wrap(~ pri_spcl_desc)+
  theme(axis.text=element_text(size=7),
        strip.text = element_text(size = 7))+
  labs(title = "Distribution of RX amount across prescribing conditions",
       subtitle = "OXAZOLIDINONES")

RX %>%
filter(usc_lvl3_desc == "OXAZOLIDINONES") %>%
  group_by(age_cat) %>%
  mutate(indivdual = n()) %>%
  ungroup() %>%
  filter(indivdual >3) %>%
  ggplot(aes(x=age_cat, y=rx_dosage_amt, color=dosage_form_nm))+
  geom_boxplot(width=0.5)+
  facet_wrap(~ unifm_prod_nm)+
  theme(axis.text=element_text(size=7),
        strip.text = element_text(size = 7))+
  labs(title = "Distribution of RX amount across age groups",
       subtitle = "OXAZOLIDINONES")
```

## are there any (potentially) unusual prescribing patterns?

### !!make rx_dosage_amt comparable

```{r}
#assisted with AI
#run & hide

#step1:Classify 49 strength descriptions into one of 8 patterns
library(dplyr)
library(stringr)
library(readr)

RX <- RX %>%
  mutate(
    
    strength_type = case_when(

      str_detect(strnt_desc, "^\\d+ MG$") ~ "single_mg",

      str_detect(strnt_desc, "^\\d+ GM$") ~ "single_g",

      str_detect(strnt_desc, "^\\d+-\\d+\\.?\\d* MG$") ~ "combo_mg",

      str_detect(strnt_desc, "^\\d+ MG/\\d+ ML$") ~ "mg_per_volume",

      str_detect(strnt_desc,
                 "^\\d+-\\d+\\.?\\d* MG/\\d+ ML$") ~ "combo_per_volume",

      str_detect(strnt_desc, "^\\d+ MG/ML$") ~ "mg_per_ml",

      str_detect(strnt_desc, "^\\d+ MG/24HR$") ~ "mg_per_24hr",

      str_detect(strnt_desc,
                 "^\\d+-\\d+\\.?\\d* MG/12HR$") ~ "combo_12hr",

      TRUE ~ "other"

    )
  )

#step2: extract the (first) ingredient
## ant1_mg = strength of the first active ingredient


RX <- RX %>%
  mutate(
    ant1_mg = case_when(

      strength_type == "single_mg" ~
        as.numeric(str_extract(strnt_desc, "^\\d+\\.?\\d*")),

      strength_type == "single_g" ~
        as.numeric(str_extract(strnt_desc, "^\\d+\\.?\\d*")) * 1000,

      strength_type %in% c("combo_mg",
                           "combo_per_volume",
                           "combo_12hr") ~
        as.numeric(str_extract(strnt_desc, "^\\d+\\.?\\d*")),

      strength_type %in% c("mg_per_volume",
                           "mg_per_ml",
                           "mg_per_24hr") ~
        as.numeric(str_extract(strnt_desc, "^\\d+\\.?\\d*")),

      TRUE ~ NA_real_
    )
  )

#step3: extract the second ingredient as there are combination antibiotics

## ant2_mg = strength of the second active ingredient

RX <- RX %>%
  mutate(
    ant2_mg = case_when(

      strength_type %in% c("combo_mg",
                           "combo_per_volume",
                           "combo_12hr") ~

        as.numeric(
          str_extract(
            strnt_desc,
            "(?<=-)\\d+\\.?\\d*"
          )
        ),

      TRUE ~ NA_real_
    )
  )

#step4: extract volumes for suspensions/liquid

## volume_ml = denominator after "/"

RX <- RX %>%
  mutate(
    volume_ml = case_when(

      strength_type %in% c("mg_per_volume",
                           "combo_per_volume") ~

        as.numeric(
          str_extract(
            strnt_desc,
            "(?<=/)\\d+"
          )
        ),

      strength_type == "mg_per_ml" ~ 1,

      TRUE ~ NA_real_
    )
  )


#then calculate the concentrations

## mg per mL


RX <- RX %>%
  mutate(

    ant1_mg_ml = case_when(

      strength_type == "mg_per_volume" ~
        ant1_mg / volume_ml,

      strength_type == "combo_per_volume" ~
        ant1_mg / volume_ml,

      strength_type == "mg_per_ml" ~
        ant1_mg,

      TRUE ~ NA_real_
    ),

    ant2_mg_ml = case_when(

      strength_type == "combo_per_volume" ~
        ant2_mg / volume_ml,

      TRUE ~ NA_real_
    )

  )

#step5: extract release hours if it's the delayed release type
## release_hours

RX <- RX %>%
  mutate(
    release_hours = case_when(

      str_detect(strnt_desc, "24HR") ~ 24,

      str_detect(strnt_desc, "12HR") ~ 12,

      TRUE ~ NA_real_
    )
  )

#step6: calculate the estimated quantity dispensed, which makes the RX amount comparable within and between antibiotics

## Estimate the number of tablets/capsules or mL dispensed


RX <- RX %>%
  mutate(

    estimated_quantity = case_when(

      strength_type %in%
        c("single_mg",
          "single_g",
          "combo_mg") ~

        rx_dosage_amt / ant1_mg,

      strength_type == "mg_per_volume" ~

        rx_dosage_amt / ant1_mg_ml,

      strength_type == "combo_per_volume" ~

        rx_dosage_amt / ant1_mg_ml,

      strength_type == "mg_per_ml" ~

        rx_dosage_amt / ant1_mg_ml,

      TRUE ~ NA_real_

    )

  )
```

### !!find & flag

```{r}
#highest strength for each antibiotic
RX <- RX %>%
  group_by(unifm_prod_nm) %>%
  mutate(
    highest_strnt =
      ant1_mg == max(ant1_mg, na.rm = TRUE)
  ) %>%
  ungroup()


#unusually large quantity (top %5)
RX <- RX %>%
  group_by(unifm_prod_nm) %>%
  mutate(
    large_quantity =
      estimated_quantity >
      quantile(
        estimated_quantity,
        0.95,
        na.rm = TRUE
      )
  ) %>%
  ungroup()
```

### make the summary table

```{r}
summary_rx <-
RX %>%
group_by(unifm_prod_nm) %>%
summarise(
  n = n(),

  median_rx =
    median(rx_dosage_amt),

  median_quantity =
    median(estimated_quantity,na.rm = TRUE),

  p95 =
    quantile(
      estimated_quantity,
      0.95,
      na.rm = TRUE),

  max_quantity =
    max(estimated_quantity),

  prop_high_strength =
    mean(highest_strnt, na.rm = TRUE),

  prop_large_quantity =
    mean(large_quantity, na.rm = TRUE)
) %>%
  arrange(desc(prop_large_quantity), by_group=TRUE)
```

### what factors are associated with stronger formulation?

```{r}
RX_filtered <- RX%>%
  group_by(unifm_prod_nm, pri_spcl_desc) %>%
  mutate(num = n()) %>%
  ungroup() %>%
  filter(num >50)

RX_model <- RX_filtered %>%
  group_by(unifm_prod_nm) %>%
  filter(n_distinct(highest_strnt) == 2) %>%
  ungroup()

RX_model$pri_spcl_desc <- factor(RX_model$pri_spcl_desc)
RX_model$provider_st_cd <- factor(RX_model$provider_st_cd)
RX_model$dosage_form_nm <- factor(RX_model$dosage_form_nm)
RX_model$strnt_desc <- factor(RX_model$strnt_desc)
RX_model$highest_strnt <- factor(RX_model$highest_strnt)
RX_model$unifm_prod_nm <- factor(RX_model$unifm_prod_nm)


  
unique(RX_filtered$unifm_prod_nm)


RX_model$unifm_prod_nm <- relevel(RX_model$unifm_prod_nm, ref = "AMOXICILLIN")
RX_model$provider_st_cd <- relevel(RX_model$provider_st_cd, ref = "NY")
RX_model$dosage_form_nm <- relevel(RX_model$dosage_form_nm, ref = "TABLET")
RX_model$pri_spcl_desc <- relevel(RX_model$pri_spcl_desc, ref = "NURSE PRACTITIONER")


model_strength<- glm(
  highest_strnt ~
    unifm_prod_nm +
    pat_age_nbr +
    pri_spcl_desc +
    provider_st_cd +
    dosage_form_nm,
  family = binomial,
  data = RX_model
)

summary(model_strength)
exp(coef(model_strength))
```

```{r}
#for selecting ref
RX_filtered %>%
  group_by(provider_st_cd) %>%
  summarize(count=n())%>%
  arrange(desc(count))

RX_filtered %>%
  filter(unifm_prod_nm == "CLINDAMYCIN HCL")


#for testing feasibility
table(RX$highest_strnt)
prop.table(table(RX$highest_strnt))
```

### what factors are associated with large prescription quantity?

```{r}
RX_filtered <- RX%>%
  group_by(unifm_prod_nm, pri_spcl_desc) %>%
  mutate(num = n()) %>%
  ungroup() %>%
  filter(num >50)

RX_filtered$pri_spcl_desc <- factor(RX_filtered$pri_spcl_desc)
RX_filtered$provider_st_cd <- factor(RX_filtered$provider_st_cd)
RX_filtered$dosage_form_nm <- factor(RX_filtered$dosage_form_nm)

  
#unique(RX_filtered$unifm_prod_nm)

amox <- RX_filtered %>%
  filter(unifm_prod_nm == "AMOXICILLIN")

amox$pri_spcl_desc <- relevel(amox$pri_spcl_desc, ref = "NURSE PRACTITIONER")
amox$provider_st_cd <- relevel(amox$provider_st_cd, ref = "NY")
amox$dosage_form_nm <- relevel(amox$dosage_form_nm, ref = "CAPSULE")


model_amox <- glm(
  estimated_quantity ~
    pat_age_nbr +
    pri_spcl_desc +
    provider_st_cd +
    dosage_form_nm,
  family = Gamma(link = "log"),
  data = amox
)

summary(model_amox)
exp(coef(model_amox))

#In the Gamma generalized linear model with a log link, patient age was positively associated with amoxicillin prescription quantity (β = 0.0025 per year, p < 0.001), corresponding to an approximately 0.25% increase in expected quantity for each additional year of age. Compared with nurse practitioners, dentists prescribed 17.9% larger quantities (exp(β) = 1.18, p < 0.001), oral and maxillofacial surgeons prescribed 13.8% larger quantities (exp(β) = 1.14, p < 0.001), and orthopedic surgeons prescribed 38.0% larger quantities (exp(β) = 1.38, p < 0.001). Differences for emergency medicine, family medicine, internal medicine, physician assistants, and pediatrics were not statistically significant after adjustment.
```

```{r}
#for selecting ref
RX %>%
filter(unifm_prod_nm == "AMOXICILLIN") %>%
  group_by(pri_spcl_desc) %>%
  mutate(num = n()) %>%
  ungroup() %>%
  filter(num >50) %>%
  ggplot(aes(x=unifm_prod_nm, y=rx_dosage_amt, color=dosage_form_nm))+
  geom_boxplot(width=0.5)+
  #facet_wrap(~ pri_spcl_desc)+
  theme(axis.text=element_text(size=7),
        strip.text = element_text(size = 7))+
  labs(title = "Distribution of RX amount across prescribing conditions",
       subtitle = "AMOXICILLIN")

RX_filtered%>%
  filter(unifm_prod_nm == "AMOXICILLIN") %>%
  group_by(dosage_form_nm) %>%
  summarize(count=n())
```

#Copayment

## make insurance model categories!!

```{r}
copayment <- copayment %>%
  mutate(insurance_type = case_when(
           str_detect(model_typ_nm, "MEDICARE") ~ "Medicare",
           str_detect(model_typ_nm, "MEDICAID") ~ "Medicaid",
           str_detect(model_typ_cd, "HIX") ~ "Health Insurance Exchange",
           str_detect(model_typ_cd, "PPO|EPO|POS|HMO|IPA|CDHP") ~ "Commercial Plans",
           str_detect(model_typ_nm, "EMPLOYER") ~ "Employer-Sponsored",
           str_detect(model_typ_nm, "FEDERAL|STATE") ~ "Government Employee Programs",
           str_detect(model_typ_cd, "CASH|DISC|VOUCHER") ~ "Self-pay",
           str_detect(model_typ_nm, "WORKER") ~ "Workers' Comp",
           str_detect(model_typ_nm, "PBM|PHARMACY BENEFIT") ~ "Pharmacy Benefit Administration",
           TRUE ~ "Other/Unspecified"
         ))
```

```{r}
copayment%>%
  select(model_typ_cd, model_typ_nm, insurance_type) %>%
  distinct()
```

```{r}
is_outlier <- function(x) {
return(x < quantile(x, 0.25) - 1.5 * IQR(x) | x > quantile(x, 0.75) + 1.5 * IQR(x))
}
copayment <- copayment %>%
group_by(usc_lvl3_desc, age_cat) %>%
mutate(outlier = ifelse(is_outlier(copay_amt), copay_amt, as.numeric(NA)))%>%
  ungroup()
```

```{r}
copayment %>%
  filter(usc_lvl3_desc == "B-LACTAM, INCREASED ACTIVITY") %>%
  ggplot(aes(x=age_cat, y=copay_amt, color=insurance_type))+
  geom_boxplot(width=0.5)+
  facet_wrap(~ unifm_prod_nm)+
  theme(axis.text=element_text(size=7),
        strip.text = element_text(size = 7))+
  labs(title = "Distribution of copayment amounts by insurance plan model type across age groups",
       subtitle = "B-LACTAM, INCREASED ACTIVITY")

#look into the only antibiotic with copayment records
copayment %>%
  filter(usc_lvl3_desc == "B-LACTAM, INCREASED ACTIVITY",
         outlier != is.na(outlier)) %>%
  group_by(pri_spcl_desc) %>%
  mutate(case=n())%>%
  filter(case>5)%>%
  ggplot(aes(x=age_cat, fill=insurance_type))+
  geom_bar()+
  facet_wrap(~pri_spcl_desc)+
  labs(title = "Distribution of copayment outliers by prescription conditions across age groups",
    subtitle="AMOXICILLIN & POT CLAVULANATE")

```

```{r}
copayment %>%
  group_by(unifm_prod_nm) %>%
  mutate(num=n())%>%
  filter(num>10)%>%
  filter(usc_lvl3_desc == "ANTNFCT, BROAD/MEDIUM SPECTRUM") %>%
  ggplot(aes(x=age_cat, y=copay_amt, color=insurance_type))+
  geom_boxplot(width=0.5)+
  facet_wrap(~ unifm_prod_nm)+
  theme(axis.text=element_text(size=7),
        strip.text = element_text(size = 7))+
  labs(title = "Distribution of copayment amounts by insurance plan model type across age groups",
       subtitle = "ANTNFCT, BROAD/MEDIUM SPECTRUM")

#look into 5 antibiotics
copayment %>%
  filter(usc_lvl3_desc == "ANTNFCT, BROAD/MEDIUM SPECTRUM",
         unifm_prod_nm == "DOXYCYCLINE (ROSACEA)",
         outlier != is.na(outlier)) %>%
  group_by(pri_spcl_desc) %>%
  mutate(case=n())%>%
  filter(case>5)%>%
  ggplot(aes(x=age_cat, fill=insurance_type))+
  geom_bar()+
  facet_wrap(~pri_spcl_desc )+
  labs(title = "Distribution of copayment outliers by prescription conditions across age groups",
    subtitle="DOXYCYCLINE (ROSACEA)")

copayment %>%
  filter(usc_lvl3_desc == "ANTNFCT, BROAD/MEDIUM SPECTRUM",
         unifm_prod_nm == "DOXYCYCLINE HYCLATE",
         outlier != is.na(outlier)) %>%
  group_by(pri_spcl_desc) %>%
  mutate(case=n())%>%
  filter(case>5)%>%
  ggplot(aes(x=age_cat, fill=insurance_type))+
  geom_bar()+
  facet_wrap(~pri_spcl_desc )+
  labs(title = "Distribution of copayment outliers by prescription conditions across age groups",
    subtitle="DOXYCYCLINE HYCLATE")


copayment %>%
  filter(usc_lvl3_desc == "ANTNFCT, BROAD/MEDIUM SPECTRUM",
         unifm_prod_nm == "MINOCYCLINE HCL",
         outlier != is.na(outlier)) %>%
  group_by(pri_spcl_desc) %>%
  mutate(case=n())%>%
  filter(case>5)%>%
  ggplot(aes(x=age_cat, fill=insurance_type))+
  geom_bar()+
  facet_wrap(~pri_spcl_desc )+
  labs(title = "Distribution of copayment outliers by prescription conditions across age groups",
    subtitle="MINOCYCLINE HCL")

copayment %>%
  filter(usc_lvl3_desc == "ANTNFCT, BROAD/MEDIUM SPECTRUM",
         unifm_prod_nm == "SARECYCLINE HCL",
         outlier != is.na(outlier)) %>%
  group_by(pri_spcl_desc) %>%
  mutate(case=n())%>%
  filter(case>5)%>%
  ggplot(aes(x=age_cat, fill=insurance_type))+
  geom_bar()+
  facet_wrap(~pri_spcl_desc )+
  labs(title = "Distribution of copayment outliers by prescription conditions across age groups",
    subtitle="SARECYCLINE HCL")

copayment %>%
  filter(usc_lvl3_desc == "ANTNFCT, BROAD/MEDIUM SPECTRUM",
         unifm_prod_nm == "VANCOMYCIN HCL",
         outlier != is.na(outlier)) %>%
  group_by(pri_spcl_desc) %>%
  mutate(case=n())%>%
  filter(case>5)%>%
  ggplot(aes(x=age_cat, fill=insurance_type))+
  geom_bar()+
  facet_wrap(~pri_spcl_desc )+
  labs(title = "Distribution of copayment outliers by prescription conditions across age groups",
    subtitle="VANCOMYCIN HCL")

```

```{r}
copayment %>%
  group_by(unifm_prod_nm) %>%
  mutate(num=n())%>%
  filter(num>10)%>%
  filter(usc_lvl3_desc == "SULFONAMIDES & TRIMETHOPRIM") %>%
  ggplot(aes(x=age_cat, y=copay_amt, color=insurance_type))+
  geom_boxplot(width=0.5)+
  facet_wrap(~ unifm_prod_nm)+
  theme(axis.text=element_text(size=7),
        strip.text = element_text(size = 7))+
  labs(title = "Distribution of copayment amounts by insurance plan model type across age groups",
       subtitle = "SULFONAMIDES & TRIMETHOPRIM")

copayment %>%
  filter(usc_lvl3_desc == "SULFONAMIDES & TRIMETHOPRIM",
         unifm_prod_nm == "SULFAMETHOXAZOLE-TRIMETHOPRIM",
         outlier != is.na(outlier)) %>%
  group_by(pri_spcl_desc) %>%
  mutate(case=n())%>%
  filter(case>5)%>%
  ggplot(aes(x=age_cat, fill=insurance_type))+
  geom_bar()+
  facet_wrap(~pri_spcl_desc )+
  labs(title = "Distribution of copayment outliers by prescription conditions across age groups",
    subtitle="SULFAMETHOXAZOLE-TRIMETHOPRIM")

```

```{r}
copayment %>%
  group_by(unifm_prod_nm) %>%
  mutate(num=n())%>%
  filter(num>10)%>%
  filter(usc_lvl3_desc == "OPHTHALMIC ANTI-INFECTIVES") %>%
  ggplot(aes(x=age_cat, y=copay_amt, color=insurance_type))+
  geom_boxplot(width=0.5)+
  facet_wrap(~ unifm_prod_nm)+
  theme(axis.text=element_text(size=7),
        strip.text = element_text(size = 7))+
  labs(title = "Distribution of copayment amounts by insurance plan model type across age groups",
       subtitle = "OPHTHALMIC ANTI-INFECTIVES")

copayment %>%
  filter(usc_lvl3_desc == "OPHTHALMIC ANTI-INFECTIVES",
         unifm_prod_nm == "OFLOXACIN (OPHTH)",
         outlier != is.na(outlier)) %>%
  group_by(pri_spcl_desc) %>%
  mutate(case=n())%>%
  filter(case>5)%>%
  ggplot(aes(x=age_cat, fill=insurance_type))+
  geom_bar()+
  facet_wrap(~pri_spcl_desc )+
  labs(title = "Distribution of copayment outliers by prescription conditions across age groups",
    subtitle="OFLOXACIN (OPHTH)")

```

```{r}
copayment %>%
  group_by(unifm_prod_nm) %>%
  mutate(num=n())%>%
  filter(num>10)%>%
  filter(usc_lvl3_desc == "OTIC ANTI-INFECTIVES") %>%
  ggplot(aes(x=age_cat, y=copay_amt, color=insurance_type))+
  geom_boxplot(width=0.5)+
  facet_wrap(~ unifm_prod_nm)+
  theme(axis.text=element_text(size=7),
        strip.text = element_text(size = 7))+
  labs(title = "Distribution of copayment amounts by insurance plan model type across age groups",
       subtitle = "OTIC ANTI-INFECTIVES")

```

```{r}
copayment %>%
  group_by(unifm_prod_nm) %>%
  mutate(num=n())%>%
  filter(num>10)%>%
  filter(usc_lvl3_desc == "VAGINAL ANTI-INFECTIVES") %>%
  ggplot(aes(x=age_cat, y=copay_amt, color=insurance_type))+
  geom_boxplot(width=0.5)+
  facet_wrap(~ unifm_prod_nm)+
  theme(axis.text=element_text(size=7),
        strip.text = element_text(size = 7))+
  labs(title = "Distribution of copayment amounts by insurance plan model type across age groups",
       subtitle = "VAGINAL ANTI-INFECTIVES")

```

```{r}
copayment %>%
  group_by(unifm_prod_nm) %>%
  mutate(num=n())%>%
  filter(num>5)%>%
  filter(usc_lvl3_desc == "ANTNFCT SYSTEMIC, OTHER") %>%
  ggplot(aes(x=age_cat, y=copay_amt, color=insurance_type))+
  geom_boxplot(width=0.5)+
  facet_wrap(~ unifm_prod_nm)+
  theme(axis.text=element_text(size=7),
        strip.text = element_text(size = 7))+
  labs(title = "Distribution of copayment amounts by insurance plan model type across age groups",
       subtitle = "ANTNFCT SYSTEMIC, OTHER")
```

```{r}
copayment %>%
  group_by(unifm_prod_nm) %>%
  mutate(num=n())%>%
  filter(num>5)%>%
  filter(usc_lvl3_desc == "OXAZOLIDINONES") %>%
  ggplot(aes(x=age_cat, y=copay_amt, color=insurance_type))+
  geom_boxplot(width=0.5)+
  facet_wrap(~ unifm_prod_nm)+
  theme(axis.text=element_text(size=7),
        strip.text = element_text(size = 7))+
  labs(title = "Distribution of copayment amounts by insurance plan model type across age groups",
       subtitle = "OXAZOLIDINONES")
```
