---
title: "IQVIA_AWaRe"
format: html
editor: visual
---

# WHO AWaRe

WHO introduced the AWaRe classification as part of the 2017 essential medicines list, as an antimicrobial stewardship tool to assist countries, territories and areas (CTAs) in their efforts toward optimizing antimicrobial use. Antibiotics are classified into three groups, Access, Watch and Reserve, considering the impact of different antibiotics and antibiotic classes on antimicrobial resistance, to emphasize the importance of their appropriate use.

```{r}
library(dplyr)
library(readr)
library(readxl)
library(ggrepel)
library(ggplot2)
library(tidyverse)
library(DHARMa)
library(scales)
```

# !!import IQVIA data

```{r}
#RX needed for extracting prescription records
unfiltered <- read_csv("Q.unfiltered100k_1.csv")
unfiltered <- unfiltered %>%
filter (pat_age_nbr >=0) %>%
  mutate(age_cat = case_when(
  pat_age_nbr < 18 ~ "Pediatrics",
  pat_age_nbr >= 65 ~ "Elders",
  TRUE ~ "Adults"
  ))

RX <- unfiltered %>%
  select(month_id,
         usc_lvl3_desc,
      unifm_prod_nm,
     pat_age_nbr,
      rx_dosage_amt,
     dosage_form_nm,
      strnt_desc,
      pri_spcl_desc,
      provider_st_cd,
      age_cat) %>%
  filter(rx_dosage_amt > 0)

```

```{r}
patient_demographics <- read_csv("Q.patient_demographics.csv")

patient_demographics <- patient_demographics %>%
  filter (pat_age_nbr >=0) %>%
  mutate(age_cat = case_when(
  pat_age_nbr < 18 ~ "Pediatrics",
  pat_age_nbr >= 65 ~ "Elders",
  TRUE ~ "Adults"
  ))
###

usage_full <- read_csv("Q.usage_with age.csv")
usage_full <- usage_full %>%
  filter_at(vars(antibiotic_name, patient_age, provider_description), all_vars(!is.na(.)))

usage_full <- usage_full %>%
  filter (patient_age >=0) %>%
  mutate(age_cat = case_when(
  patient_age < 18 ~ "Pediatrics",
  patient_age >= 65 ~ "Elders",
  TRUE ~ "Adults"
  ))
```

# make top used antibiotic groups within each category in IQVIA data

```{r}
patient_demographics %>%
  group_by(usc_lvl3_desc) %>%
  summarise(n = n_distinct(unifm_prod_nm))%>%
  arrange(desc(n))

top_prod <- patient_demographics %>%
  filter(usc_lvl3_desc == "ANTNFCT SYSTEMIC, OTHER") %>%
  group_by(unifm_prod_nm) %>%
  summarise(count = sum(n)) %>%
  arrange(desc(count)) %>%
  slice_head(n=5) 
antnfct_systemic<-top_prod$unifm_prod_nm


top_prod <- patient_demographics %>%
  filter(usc_lvl3_desc == "ANTNFCT, BROAD/MEDIUM SPECTRUM") %>%
  group_by(unifm_prod_nm) %>%
  summarise(count = sum(n)) %>%
  arrange(desc(count)) %>%
  slice_head(n=5) 
antnfct_broad<-top_prod$unifm_prod_nm


top_prod <- patient_demographics %>%
  filter(usc_lvl3_desc == "B-LACTAM, INCREASED ACTIVITY") %>%
  group_by(unifm_prod_nm) %>%
  summarise(count = sum(n)) %>%
  arrange(desc(count)) %>%
  slice_head(n=5) 
b_lactam<-top_prod$unifm_prod_nm


top_prod <- patient_demographics %>%
  filter(usc_lvl3_desc == "OPHTHALMIC ANTI-INFECTIVES") %>%
  group_by(unifm_prod_nm) %>%
  summarise(count = sum(n)) %>%
  arrange(desc(count)) %>%
  slice_head(n=5) 
ophthalmic<-top_prod$unifm_prod_nm


top_prod <- patient_demographics %>%
  filter(usc_lvl3_desc == "OTIC ANTI-INFECTIVES") %>%
  group_by(unifm_prod_nm) %>%
  summarise(count = sum(n)) %>%
  arrange(desc(count)) %>%
  slice_head(n=5) 
otic<-top_prod$unifm_prod_nm


top_prod <- patient_demographics %>%
  filter(usc_lvl3_desc == "OXAZOLIDINONES") %>%
  group_by(unifm_prod_nm) %>%
  summarise(count = sum(n)) %>%
  arrange(desc(count)) %>%
  slice_head(n=5) 
oxazolidinones<-top_prod$unifm_prod_nm


top_prod <- patient_demographics %>%
  filter(usc_lvl3_desc == "SULFONAMIDES & TRIMETHOPRIM") %>%
  group_by(unifm_prod_nm) %>%
  summarise(count = sum(n)) %>%
  arrange(desc(count)) %>%
  slice_head(n=5) 
sulf_trim<-top_prod$unifm_prod_nm


top_prod <- patient_demographics %>%
  filter(usc_lvl3_desc == "VAGINAL ANTI-INFECTIVES") %>%
  group_by(unifm_prod_nm) %>%
  summarise(count = sum(n)) %>%
  arrange(desc(count)) %>%
  slice_head(n=5) 
vaginal<-top_prod$unifm_prod_nm
```

# !!import & clean WHO data

```{r}
WHO_cat <- read_xlsx("WHO_AWARE_EML_cat.xlsx")

WHO_cat <- WHO_cat %>%
 mutate(EML_2025 = case_when (
     EML_2025 == 'Access group antibiotics' ~ 'Access',
     EML_2025 == 'Reserve group antibiotics' ~ 'Reserve',
     EML_2025 == 'Watch group antibiotics' ~ 'Watch',
     EML_2025 == 'Antituberculosis medicines' ~ 'Antituberculosis',
     EML_2025 == 'Antileprosy medicines' ~ 'Antileprosy'
  ))

```

```{r}
#no longer useful as DDD is not available

AWARE <- read_csv("WHO_Pattern of AWaRe antibiotic use at national level_2025.csv")
AWARE_cleaned<- AWARE %>% select(Period, c(5,8), Dim1, Value)

AWARE_cleaned <- AWARE_cleaned %>% pivot_wider(
  names_from = Period,
  values_from = Value
)

AWARE_cleaned <- AWARE_cleaned %>%
   mutate(Dim1 = case_when (
     Dim1 == 'Not classified (%) / not recommended (%)' ~ 'Not_EML',
     Dim1 == 'Reserve (%)' ~ 'Reserve',
     Dim1 == 'Watch (%)' ~ 'Watch',
     Dim1 == 'Access (%)' ~ 'Access'))

AWARE_cleaned <- AWARE_cleaned %>% rename("%by_cat" = "Dim1")

```

#Add AWaRe categories to the IQVIA dataset

```{r}
WHO_catCombine <- WHO_cat %>%
  mutate(Antibiotic = toupper(Antibiotic))
```

```{r}
RX_labeled <- RX %>%
  distinct(unifm_prod_nm)

WHO_catCombine <- WHO_catCombine %>%
  mutate(key = word(Antibiotic, 1)) %>%
  select(-c(3,5,7,9))

RX_labeled <- RX_labeled %>%
  mutate(
    match = case_when(
      unifm_prod_nm == "CIPROFLOXACIN-CIPROFLOXACIN HCL" ~ "CIPROFLOXACIN",
      unifm_prod_nm == "AMOXICILLIN & POT CLAVULANATE" ~ "AMOXICILLIN/CLAVULANIC-ACID",
      unifm_prod_nm == "SULFAMETHOXAZOLE-TRIMETHOPRIM" ~ "SULFAMETHOXAZOLE/TRIMETHOPRIM",
      unifm_prod_nm == "CEPHALEXIN" ~ "CEFALEXIN",
      unifm_prod_nm == "PENICILLIN V POTASSIUM" ~ "PHENOXYMETHYLPENICILLIN",
      unifm_prod_nm == "VANCOMYCIN HCL" ~ "VANCOMYCIN_ORAL",
      unifm_prod_nm == "NEOMYCIN SULFATE" ~ "NEOMYCIN_IV",
      unifm_prod_nm == "MINOCYCLINE HCL" ~ "MINOCYCLINE_IV",
      unifm_prod_nm == "CEFPODOXIME PROXETIL" ~ "CEFPODOXIME-PROXETIL",
      TRUE ~ unifm_prod_nm))

RX_labeled <- RX_labeled %>%
  mutate(key = if_else(
  str_detect(match, " \\+ "),
  match,
  word(match, 1)
))

RX_labeled<- RX_labeled %>%
  left_join(
    WHO_catCombine,
    by = "key") %>%
  select(-c(2,3))
```

```{r}
writexl::write_xlsx(RX_labeled, "IQVIA21_RX_labeled.xlsx")
```

#Analysis

##What are the frequencies of A, W, R antibiotic prescriptions, respectively

```{r}
RX <- RX %>%
  left_join(RX_labeled %>% 
              select(c(1, 4, 5, 6, 7)),
            by = "unifm_prod_nm")

aware2025<- RX%>%count(AWARE_Cat2025) %>%
  mutate(perc_2025 = round(n / sum(n), 4)*100)
aware2021<- RX%>%count(AWARE_Cat2021) %>%
  mutate(perc_2021 = round(n / sum(n), 4)*100)
aware2019<- RX%>%count(AWARE_Cat2019) %>%
  mutate(perc_2019 = round(n / sum(n), 4)*100)
eml2025<- RX%>%count(EML_2025) %>%
  mutate(perc_eml2025 = round(n / sum(n), 4)*100)
```

```{r}
RX2018<-RX%>%
  filter(str_starts(month_id,"2018"))
RX2019<-RX%>%
  filter(str_starts(month_id,"2019"))
RX2020<-RX%>%
  filter(str_starts(month_id,"2020"))

RX2018%>%count(AWARE_Cat2025) %>%
  mutate(perc_2019 = round(n / sum(n), 4)*100)
RX2019%>%count(AWARE_Cat2025) %>%
  mutate(perc_2019 = round(n / sum(n), 4)*100)
RX2020%>%count(AWARE_Cat2025) %>%
  mutate(perc_2019 = round(n / sum(n), 4)*100)

```

```{r}
sumRX = sum(RX$rx_dosage_amt)
RX %>%
  group_by(AWARE_Cat2025) %>%
  summarise(totalRX=sum(rx_dosage_amt),
            totalRX_share = totalRX/sumRX*100)
```

```{r}
noAMOXICILLIN<-RX %>%
  filter(!str_starts(unifm_prod_nm,"AMOXICILLIN"))
sumRX = sum(noAMOXICILLIN$rx_dosage_amt)
noAMOXICILLIN%>%
  group_by(AWARE_Cat2025) %>%
  summarise(totalRX=sum(rx_dosage_amt),
            totalRX_share = totalRX/sumRX*100)
```


```{r}
#| echo: false
##remove redundant tabels
#rm(list = c("awareGlobal_A", "awareGlobal_R", "unfiltered", "WHO_catCombine", "AWARE"))
```

# 
